$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-06 paper/paper.tex
Do not number equations
Clarifications

We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system with leftovers (\autoref{leftover-typing}).
\begin{nidefinition}[Indexed set of usage algebras]
  An \emph{indexed set of usage algebras} is a nonempty type $\Idx$ of indices together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras (\autoref{fig:indexed-multiplicities}).
  \begin{figure}[h]
    \begin{equation}
      \begin{split}
        &\Idx               &:{} &\Set \\
        &\type{\exists IDX} &:{} &\Idx \\
        &\Usage             &:{} &\Idx \to \Set \\
        &\type{ALGEBRAS}    &:{} &(idx : \Idx) \to \Algebra_{\Usage_{idx}}
      \end{split}
    \end{equation}
    \caption{Indexed set of usage algebras.}
    \label{fig:indexed-multiplicities}
  \end{figure}
\end{nidefinition}

@@

Removing an unused variable preserves the well-typedness of a process.
Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
\begin{nitheorem}[Strengthening]
  \label{thm:strengthening} 
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Let $i$ be a variable index not in $P$, such that $\Unused_i \; P$.
  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
\end{nitheorem}

\begin{remark}
  Usage annotations of a variable $i$ within a process $P$ can be preserved as leftovers, yet the variable $i$ be mentioned by $P$: $P$ may send $\lz$ usage annotations of $i$ over a channel.
\end{remark}

Swapping two variables preserves the well-typedness of a process.
For simplicity, we extend $\func{swap}_i$ introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
\begin{nitheorem}[Swapping]
  \label{thm:swapping}
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
\end{nitheorem}

>>>>

We use indexed sets of usage algebras to allow several usage algebras to coexist in our type system with leftovers (\autoref{leftover-typing}).
\begin{nidefinition}[Indexed set of usage algebras]
  An \emph{indexed set of usage algebras} is a nonempty type $\Idx$ of indices together with an interpretation $\Usage$ of indices into types, and an interpretation $\type{ALGEBRAS}$ of indices into usage algebras (\autoref{fig:indexed-multiplicities}).
  \begin{figure}[h]
    \begin{equation*}
      \begin{split}
        &\Idx               &:{} &\Set \\
        &\type{\exists IDX} &:{} &\Idx \\
        &\Usage             &:{} &\Idx \to \Set \\
        &\type{ALGEBRAS}    &:{} &(idx : \Idx) \to \Algebra_{\Usage_{idx}}
      \end{split}
    \end{equation*}
    \caption{Indexed set of usage algebras.}
    \label{fig:indexed-multiplicities}
  \end{figure}
\end{nidefinition}

@@

Removing an unused variable preserves the well-typedness of a process.
Let $\func{del}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
\begin{nitheorem}[Strengthening]
  \label{thm:strengthening} 
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Let $i$ be a variable not in $P$, such that $\Unused_i \; P$.
  Then lowering every variable greater than $i$ in $P$ is well typed in $\types{\func{del}_i \; \gamma}{\func{del}_i \; \Gamma}{\func{lower}_i \; P}{\func{del}_i \; \Xi}$.
\end{nitheorem}

\begin{remark}
  The \emph{unusedness} predicate cannot depend on the preservation of usage annotations by a process: the process could preserve such annotations, yet sent a $\lz$ usage of those variables over a channel.
\end{remark}

Swapping two variables preserves the well-typedness of a process.
For simplicity, we extend $\func{swap}_i$ introduced in \autoref{structural-congruence} to swap types in typing contexts and usage annotations in usage contexts.
\begin{nitheorem}[Swapping]
  \label{thm:swapping}
  Let $P$ be well typed in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\allowbreak \func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
\end{nitheorem}
